Hierarchical Controller Architecture in Software Defined Networks

### 2013-12-14 paper/paper.tex
author list changed to length 3
Evaluation sections

\title{Hierarchical Controller Architecture in Software Defined Networks}
\author{
        Author 1 \\
        UIUC\\
            \and
        Author 2\\
        UIUC\\
}
\date{\today}

\documentclass[10pt, twocolumn]{article}

@@

\paragraph{Outline}
The remainder of this article is organized as follows. Section~\ref{sec:hierarchical} presents Hierarchical Controller Architecture. We present the experimental results in section~\ref{sec:eval}. In section~\ref{sec:timestamps}, we present an optimization to conventional controller architecture to help improve the performance of application even if states between controllers is inconsistent. Related work is presented in section~\ref{sec:related}. We discuss ............Our new and exciting results are described in Section~\ref{results}.

@@

\section{Evaluation}
\label{sec:eval}
In this section we describe the results.

@@

\end{document}
This is never printed

>>>>

\title{Hierarchical Controller Architecture in Software Defined Networks}
\author{
        Author 1 \\
            \and
        Author 2\\
        %UIUC\\
			\and
		Author 3\\
}
\date{\today}

\documentclass[10pt, twocolumn]{article}
\usepackage{url,graphicx,tabularx,array,geometry, caption, subcaption, float}

%-- Commands for header
\renewcommand{\title}[1]{\textbf{#1}\\}
\renewcommand{\line}{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}\hline\\\end{tabularx}\\[-0.10cm]}
\newcommand{\leftright}[2]{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}#1%
& #2\\\end{tabularx}\\[-0.10cm]}

\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

@@

\paragraph{Outline}
The remainder of this article is organized as follows. Section~\ref{sec:hierarchical} presents Hierarchical Controller Architecture. We present the experimental results in section~\ref{sec:eval}. In section~\ref{sec:timestamps}, we present an optimization to conventional controller architecture to help improve the performance of application even if states between controllers is inconsistent. Related work is presented in section~\ref{sec:related}. We discuss ............Our new and exciting results are described in Section~\ref{sec:eval}.

@@

\section{Evaluation}
\label{sec:eval}
We simulated a dummy control application on both hierarchical as well as flat controller architecture, with same network topology and flow arrival distribution. The details of simulation, metrics and results are presented in this section.

\subsection{Implementation}
\label{sec:implementation}
We wrote our own simulator, which is available at [TODO REF simulator]. We did not use an existing simulator/controller such as Mininet with Floodlight [TODO ref] for the following two reasons: (1) Existing controller lacks the support for flexible controller architecture and are not suitable for hierarchical arrangement of controllers. Modifying existing controllers such as Floodlight to add support for One-switch abstraction (section~\ref{subsec:aggr}), is not a good idea because it requires code/design changes in their primitive modules. Besides, we were not looking for a feature rich controller at that point of time. (2) Existing simulators do not provide controls to regulate replication between controllers or to measure communication overhead (number of messages exchanged ) between controllers, which we needed for evaluation of control plane architecture. Given the above requirements, it seemed easier to write a simulator which provides one-switch abstraction support for the controllers. There are three prime interfaces: (1) {\it Switch Data Plane:} It is implemented by a switch to create forwarding elements to handle packet forwarding according to forwarding rules. (2) {\it Switch Control Plane:} It is implemented by both controllers and switches and is used to establish connections to parent controllers to send packets which do not match any forwarding rule and to receive forwarding rules from parent controller. (3) {\it Controller Control Plane:} It is implemented by controllers to establish connections to children switches/controllers. It is used to receive packets which do not match any forwarding rule and to send forwarding rules to children switches/controllers.Simulator defines two networks: (1) {\t Data Network:} This network consists of all switches (and no controller) connected through links. (2) {\it Management Network:} This network consists of all controllers and switches connected through Controller - Switch Control Plane interfaces. This network allows flexible controller architecture, which could be hierarchical or flat.

To observe the effect of different architecture in real world applications, we used SWIM project [TODO ref swim], which provides synthetic representative test workloads  by sampling historical MapReduce cluster traces using [swim paper ref TODO]. The original trace comes from historical Hadoop traces on a 600 machine cluster at Facebook and spans 6 months, and contains roughly 1 million jobs. The representative trace file contains ~6000 jobs. The trace was simulated with a cluster of 600 machines (as in the original cluster) connected through a network of 200 switches and 10 Gbps links. The cluster is divided into 10 control domains, with 60 hosts, 20 switches and one controller for each domain. Each switch in a domain is connected to 1 to 4 switches within the same domain, with degree being chosen randomly. Each domain contains two Top-of-the-rack switches which inter-connects domains. For hierarchical architecture, there are 11 controllers: 1 controller per domain and one root controllers which manages the other 10 controllers. While flat architecture consists of 10 peer controllers, one or each domain.

Control application is a TE-cum-routing application. When a new flow arrives, control application finds the path with minimum maximum utilization and sets up forwarding rules (based on IP destination and incoming ports) along the path. If multiple such paths exists (which is frequently the case, due to big network), shortest path (in term of number of hops) is chosen. This application runs on each controller in both hierarchical and flat architecture. However, in the hierarchical architecture, the view of application is limited and may consists of controllers or switches as forwarding elements, while in the flat architecture, control application is presented with global view of the network. For simplicity, we have kept network topology as static i.e. links/nodes do not fail and/or created during the simulation. For each Hadoop job, a dummy MapReduce application choses directory, map and reduce nodes randomly in the network, thus, creating a number of flows which are handled by Control application running on controllers. The duration of flow depends on the bytes needed to transfer and assigned rates.

\subsection{Results}
\label{subsec:results}

\begin{figure}[t]
\centering
\begin{subfigure}{.5\textwidth}
\label{fig:path}
\includegraphics[scale=0.3]{path}
\caption{Application performance: path length for flows}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
\label{fig:util}
\includegraphics[scale=0.3]{util}
\caption{Application performance: Maximum link utilization for flows}
\end{subfigure}
\caption{Application performance}
\end{figure}

\begin{figure}
\label{fig:msg}
\includegraphics[scale=0.3]{msg}
\caption{Communication overhead per flow}
\end{figure}

We simulated the trace file for both flat and hierarchical architecture, with same topology and same assignment of task nodes for each job. We calculated two metrics: (1) The performance of control application which is measured as the length path (number of hops) assigned to flows (Figure ~\ref{fig:path}) and maximum link utilization encountered by the flow (Figure ~\ref{fig:util}). While the flat architecture gives optimal results, hierarchical architecture provides sub-optimal results. (2) Communication overhead during flow set up, which is measured in term of number of messages passed between switches and controllers as well as within controllers (Figure ~\ref{fig:msg}). The simulation assumes instant replication of network state between controllers. Results show that hierarchical architecture reduces the communication overhead between controllers by an order of magnitude while modestly deviating from optimal solution.

\begin{table*}
\centering
\begin{tabular}{| L{5cm} | L{3cm} | L{3cm} | L{5cm} |}
\hline
& {Average path length} & {Average maximum utilization} & {Average communication overhead per flow} \\ \hline
Flat architecture & 7.81 & 24.62 & 153.54 \\ \hline
Hierarchical architecture & 9.17 & 27.78 & 14.52 \\ \hline
\end{tabular}
\end{table*}%

@@

\end{document}
